\documentclass[12pt]{article}
\usepackage{amssymb, amsmath, array}
\usepackage[xdvi]{graphics,color}
\usepackage[xdvi]{graphicx}
\usepackage{lscape}
\usepackage{pdflscape}
%\usepackage{xspace}
%\usepackage{enumerate}
\usepackage{threeparttable}
\usepackage{latexsym}

%\setlength{\textwidth}{18cm} \setlength{\textheight}{22.5cm}
%\setlength{\headheight}{-1cm} \setlength{\footskip}{1cm}
%\setlength{\oddsidemargin}{-1cm} \setlength{\evensidemargin}{0cm}
\setlength{\textwidth}{17.1cm} \setlength{\textheight}{22.5cm}
\setlength{\headheight}{-1cm} \setlength{\footskip}{1cm}
\setlength{\oddsidemargin}{-0.5cm}
\setlength{\evensidemargin}{-0.5cm}
\begin{document}
\baselineskip 20pt
\newtheorem{proposition}{Proposition}
\newtheorem{assumption}{Technical Assumption}

\begin{titlepage}
\title{{Labor Market Hours Constraints and Home Production}\thanks{We thank Daniel Aaronson, Charlie Brown, Karen Dynan, Dmitriy Stolyarov, and seminar and conference participants for helpful discussions and comments. Colin Motley, Hannah Farkas, and Ilaria Malisan provided superb research assistance. The views presented in this paper are those of the author and are not necessarily those of the Federal Reserve Board and its staff.}
\\ \vspace{0.5in}
}
  \author{Geng Li\thanks{Federal Reserve Board, Washington, DC 20551.  E-mail: Geng.Li@frb.gov. }  \\
          Federal Reserve Board
          \and Brett McCully\thanks{Collegio Carlo Alberto, Turin, Italy 10122.  E-mail: brett.mccully@carloalberto.org.} \\
          Collegio Carlo Alberto\vspace{.6in}}
\bigskip
\date{\today}
\maketitle\thispagestyle{empty}
\begin{abstract}
We document that labor market hours constraints are more pervasive than suggested by standard measures of labor market slack. We show that households adjust to such market hours constraints by increasing time spent on home production. 
\end{abstract}
\end{titlepage}
\clearpage

\input{intro.tex}
\input{data.tex}
\input{trends.tex}
\input{response.tex}
\input{conclusion.tex}
******Cite Baker and Yannelis
\input{tables.tex}

\end{document}

